---
title: "Denitrifying (N) Gene Presence in Milled vs Unmilled Woodchips"
author: "Katie L. Duggan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Files needed:

- "Woodchip_readVrefs_N.csv" (output from DIAMOND - copied readVrefs file into csv)
- "annotation_map.csv" (created by Jim for denitrifying genes)
- "Woodchip_metadata_N.csv" (total read counts for each sample; read from KBase)

## Load packages & set directory
```{r, include=FALSE}
library(tidyverse)

setwd("C:/github/WBR-microbiome")
```

## Making the plot

### Read in the DIAMOND output data
```{r}
dmnd_out_Ncycle = read_csv("C:/github/WBR-microbiome/data/Woodchip_readVrefs_N.csv", 
                  col_types = "ccdiiiiiiiddc", 
                  col_names = c("query", 
                                "subject", 
                                "pid",
                                "aln_len", 
                                "mismatches", 
                                "gaps", 
                                "qstart", 
                                "qend", 
                                "sstart", 
                                "send", 
                                "e-value", 
                                "bit", 
                                "fn")) %>%
    filter(pid >= 60, aln_len > 25)
```

### Correct the sample IDs
```{r}
dmnd_out_Ncycle_correct <- dmnd_out_Ncycle %>% 
  mutate(fn = ifelse(fn == "./7G.out.daa", "sample7G", fn)) %>% 
  mutate(fn = ifelse(fn == "./8W.out.daa", "sample8W", fn)) %>%
  mutate(fn = ifelse(fn == "./9G.out.daa", "sample9G", fn)) %>%
  mutate(fn = ifelse(fn == "./10W.out.daa", "sample10W", fn)) %>%
  mutate(fn = ifelse(fn == "./11G.out.daa", "sample11G", fn)) %>%
  mutate(fn = ifelse(fn == "./12W.out.daa", "sample12W", fn))
```

### Read in the annotation map for denitrifying genes
```{r}
annotation_map = read_csv("C:/github/WBR-microbiome/data/annotation_map.csv", col_types = "cc", col_names = c("gene","subject"))
```

### Read in the metadata (i.e., read counts)
```{r}
read_count_N = read_csv("C:/github/WBR-microbiome/data/Woodchip_metadata_N.csv") %>%
  rename(fn = filename)
```

### Join the three dataframes together
```{r}
dmnd_N = dmnd_out_Ncycle_correct %>%
    left_join(annotation_map) %>%
    group_by(fn, gene) %>%
    summarize(raw.count = n()) %>%
    left_join(read_count_N) %>%
    mutate(norm.count = (raw.count / count) * 1000000) #calculate counts
```

### Add gene names
```{r}
d2_N = dmnd_N[dmnd_N$gene %in% c("Nap", "Nar", "NirK", "NirS", "cNorB", "qNorZ", "NosZ"), ] 
#filters out any rows that are not these 7 genes
```

### Group by treatment (i.e., "G" for milled, "W" for unmilled)
```{r}
d3_N = dplyr::group_by(d2_N,gene, Treatment)

d3_N_avg = dplyr::summarize(d3_N, mean = mean(norm.count), sd = sd(norm.count))
```

### Add a new column with corrected names
```{r}
d3_N_avg$rename = c("cNor", "cNor","Nap", "Nap", "Nar", "Nar", "NirK", "NirK", "NirS", "NirS", "NosZ", "NosZ", "qNor", "qNor") 
```

### Put genes in order for the facet
```{r}
d3_N_avg$ord = factor(d3_N_avg$rename, levels=c("Nap","Nar","NirK","NirS","cNor","qNor","NosZ"))
```

### Create the dot chart
```{r}
p = ggplot(d3_N_avg, aes(mean, Treatment)) + geom_point()

p = p +  geom_errorbarh(aes(xmax = mean + sd, xmin = mean - sd, height = .5)) + expand_limits(x = 0)

p = p + facet_grid(ord ~ .)

p = p + theme(panel.background = element_rect(fill = "white"), panel.grid.major.x = element_blank(), strip.background =element_rect(fill="gold"), panel.grid.major.y = element_line(linetype=3, color="dark grey"), legend.position="none", axis.text.y = element_text(size = 10)) #add legend position none.

p = p + labs(x = "Reads per million total reads", y = "Treatment")
  
p
```

### Find dimensions of plot in inches
```{r}
dev.size("in")
```

### Plot with appropriate dimensions
```{r}
tiff("C:/github/WBR-microbiome/figures/N_presence_dotchart.tiff", width = 8.4, height = 4, units = 'in', res = 150)
plot(p) # Make plot
dev.off()
```

