---
title: "Polyphosphate kinase (ppk) Gene Presence in Milled vs Unmilled Woodchips"
author: "Katie L. Duggan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Files needed:

- "Woodchip_readVrefs_ppk.csv" (output from DIAMOND - copied readVrefs file into csv)
- "Woodchip_metadata_ppk.csv" (total read counts for each sample; read from KBase)

## Load packages & set directory
```{r, include=FALSE}
library(tidyverse)

setwd("C:/github/WBR-microbiome")
```

## Making the plot

### Read in the DIAMOND output data
```{r}
dia_ppk.df = read_csv("C:/github/WBR-microbiome/data/Woodchip_readVrefs_ppk.csv", 
                  col_types = "ccdiiiiiiiddc", 
                  col_names = c("query", 
                                "subject", 
                                "pid",
                                "aln_len", 
                                "mismatches", 
                                "gaps", 
                                "qstart", 
                                "qend,", 
                                "sstart", 
                                "send", 
                                "e-value", 
                                "bit", 
                                "fn")) %>%
    filter(pid >= 60, aln_len > 25)
```

### Read in the metadata (i.e., read counts)
```{r}
count.df = read_csv("C:/github/WBR-microbiome/data/Woodchip_metadata_ppk.csv") %>%
  rename(fn = filename)
```

### Join the two dataframes together
```{r}
d = dia_ppk.df %>%
    group_by(fn) %>%
    summarize(raw.count = n()) %>%
    left_join(count.df) %>%
    mutate(norm.count = (raw.count / count) * 1000000)
```


### Group by treatment (i.e., "G" for milled, "W" for unmilled)
```{r}
d3 = dplyr::group_by(d, Treatment)

d3.avg = dplyr::summarize(d3, mean = mean(norm.count), sd = sd(norm.count))
```


### Remove gene labels on right
```{r}
p = ggplot(d3.avg, aes(mean, Treatment)) + geom_point()

p = p +  geom_errorbarh(aes(xmax = mean + sd, xmin = mean - sd, height = .5))+ expand_limits(x = 0)

p = p + theme(panel.background = element_rect(fill = "white"), panel.grid.major.x = element_blank(), strip.background =element_rect(fill="gold"), panel.grid.major.y = element_line(linetype=3, color="dark grey"), legend.position="none", axis.text.y = element_text(size = 5)) #add legend position none.

p = p + labs(x = "Reads per million total reads", y = "Treatment")
  
p
```

### Find dimensions of plot in inches
```{r}
dev.size("in")
```

### Plot with appropriate dimensions
```{r}
tiff("C:/github/WBR-microbiome/figures/ppk_presence_dotchart.tiff", width = 8.4, height = 3, units = 'in', res = 150)
plot(p) # Make plot
dev.off()
```

