---
title: "Carbohydrate-Active enZyme (CAZy) Presence in Milled vs Unmilled Woodchips"
author: "Katie L. Duggan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Files needed:

- "readVrefs_CAZy.m8" (output from DIAMOND - too large to copy into a csv)
- "Woodchip_metadata_CAZy.csv" (total read counts for each sample; read from KBase)

## Load packages & set directory
```{r, include=FALSE}
library(tidyverse)

setwd("C:/github/WBR-microbiome")
```

## Making the plot

### Read in the DIAMOND output data
```{r}
dmnd_out_CAZy <- read_delim(file = "C:/github/WBR-microbiome/data/readVrefs_CAZy.m8")

dmnd_out_CAZy <- dmnd_out_CAZy %>% 
  mutate(source = gsub("\\|.*", "", subject),
         temp = gsub("^[^\\|]+\\||-.*", "", subject),
         classI = gsub("\\|.*", "", temp),
         classII = sub(".*\\|", "", temp))
```

### Correct the sample IDs
```{r}
dmnd_out_CAZy <- dmnd_out_CAZy %>% 
  mutate(fn = ifelse(fn == "./7G.out.daa", "sample7G", fn)) %>% 
  mutate(fn = ifelse(fn == "./8W.out.daa", "sample8W", fn)) %>%
  mutate(fn = ifelse(fn == "./9G.out.daa", "sample9G", fn)) %>%
  mutate(fn = ifelse(fn == "./10W.out.daa", "sample10W", fn)) %>%
  mutate(fn = ifelse(fn == "./11G.out.daa", "sample11G", fn)) %>%
  mutate(fn = ifelse(fn == "./12W.out.daa", "sample12W", fn))
```

### Read in the metadata (i.e., read counts)
```{r}
read_count_CAZy <- read_csv("C:/github/WBR-microbiome/data/Woodchip_metadata_CAZy.csv", col_types = cols(filename = col_character(), Treatment = col_character(), count = col_double())) %>%
  dplyr::rename(fn = filename)
```

### Join the two dataframes together
```{r}
dmnd_CAZy <- dmnd_out_CAZy %>%
  left_join(read_count_CAZy) %>%
  select(query, fn, count, classI)
```

### Gives a count of the AA types
```{r}
dmnd_AAcount <- dmnd_CAZy %>%
  count(classI) %>% 
  filter(classI == "AA0" | classI == "AA1" | classI == "AA1_1" | classI == "AA1_2" | classI == "AA1_3" | classI == "AA2" | classI == "AA3" | classI == "AA3_1" | classI == "AA3_2" | classI == "AA3_3" | classI == "AA4" | classI == "AA5" | classI == "AA5_1" | classI == "AA5_2" | classI == "AA6" | classI == "AA7" | classI == "AA8" | classI == "AA9" | classI == "AA10" | classI == "AA11" | classI == "AA12" | classI == "AA13" | classI == "AA14" | classI == "AA15" | classI == "AA16" | classI == "AA17")
```

### Filter based on AA
```{r}
dmnd_AAfilter <- dmnd_CAZy %>% 
  filter(classI == "AA0" | classI == "AA1" | classI == "AA1_1" | classI == "AA1_2" | classI == "AA1_3" | classI == "AA2" | classI == "AA3" | classI == "AA3_1" | classI == "AA3_2" | classI == "AA3_3" | classI == "AA4" | classI == "AA5" | classI == "AA5_1" | classI == "AA5_2" | classI == "AA6" | classI == "AA7" | classI == "AA8" | classI == "AA9" | classI == "AA10" | classI == "AA11" | classI == "AA12" | classI == "AA13" | classI == "AA14" | classI == "AA15" | classI == "AA16" | classI == "AA17")
```

### Summarize by sample ID
```{r}
d2_CAZy <- dmnd_AAfilter %>%
    group_by(fn, classI) %>%
    summarize(raw.count = n()) %>% #gives number of queries in each sample were mapped to the given gene
    left_join(read_count_CAZy) %>%
    mutate(norm.count = (raw.count / count) * 1000000) #gives reads per million total reads
```

### Group by treatment (i.e., "G" for milled, "W" for unmilled)
```{r}
d3_CAZy = dplyr::group_by(d2_CAZy, classI, Treatment)

d3_CAZy_avg = dplyr::summarize(d3_CAZy, mean = mean(norm.count), sd = sd(norm.count))
```

### Add a new column with corrected names
```{r}
d3_CAZy_avg$rename = c("AA0", "AA0", "AA1", "AA1", "AA1_1", "AA1_1", "AA1_2", "AA1_2", "AA1_3", "AA1_3", "AA10", "AA10", "AA11", "AA11", "AA12", "AA12", "AA13", "AA13", "AA14", "AA14", "AA15", "AA15", "AA16", "AA16", "AA17", "AA17",  "AA2", "AA2", "AA3", "AA3", "AA3_1", "AA3_1", "AA3_2", "AA3_2", "AA3_3", "AA3_3", "AA4", "AA4", "AA5", "AA5", "AA5_1", "AA5_1", "AA5_2", "AA5_2", "AA6", "AA6", "AA7", "AA7", "AA8", "AA8", "AA9", "AA9")

#adds a column of these values, in this order (FYI - use d3d3_CAZy_avg to help make this list in the correct order)
```

### Put CAZy AA families in order for the facet
```{r}
d3_CAZy_avg$ord = factor(d3_CAZy_avg$rename, levels=c("AA0", "AA1", "AA1_1", "AA1_2", "AA1_3", "AA2", "AA3", "AA3_1", "AA3_2", "AA3_3", "AA4", "AA5", "AA5_1", "AA5_2", "AA6", "AA7", "AA8", "AA9", "AA10", "AA11", "AA12", "AA13", "AA14", "AA15", "AA16", "AA17"))
```

### Create the dot chart
```{r}
#p = ggplot(d3_CAZy_avg, aes(mean, Treatment)) + geom_point()

#p = p +  geom_errorbarh(aes(xmax = mean + sd, xmin = mean - sd, height = .5))

#p = p + facet_grid(ord ~ .)

#p = p + theme(panel.background = element_rect(fill = "white"), panel.grid.major.x = element_blank(), strip.background =element_rect(fill="gold"), panel.grid.major.y = element_line(linetype=3, color="dark grey"), legend.position="none", axis.text.y = element_text(size = 5)) #add legend position none.

#p = p + labs(x = "Reads per million total reads", y = "Treatment")
  
#p
```

### Find dimensions of plot in inches
```{r}
#dev.size("in")
```

### Plot with appropriate dimensions
```{r}
#tiff("C:/github/WBR-microbiome/figures/CAZy_presence_dotchart.tiff", width = 8.5, height = 11, units = 'in', res = 150)
#plot(p) # Make plot
#dev.off()
```

### Trim/subset the plot to exclude AA families with low gene presence, and create the dot chart
```{r}
d3_CAZy_avg_over15 <- d3_CAZy_avg %>% 
  filter(mean > 15)

p2 = ggplot(d3_CAZy_avg_over15, aes(mean, Treatment)) + geom_point()

p2 = p2 +  geom_errorbarh(aes(xmax = mean + sd, xmin = mean - sd, height = .5)) + expand_limits(x = 0)

p2 = p2 + facet_grid(ord ~ .)

p2 = p2 + theme(panel.background = element_rect(fill = "white"), panel.grid.major.x = element_blank(), strip.background =element_rect(fill="gold"), panel.grid.major.y = element_line(linetype=3, color="dark grey"), legend.position="none", axis.text.y = element_text(size = 10)) #add legend position none.

p2 = p2 + labs(x = "Reads per million total reads", y = "Treatment")
  
p2
```

### Find dimensions of plot in inches
```{r}
dev.size("in")
```

### Plot with appropriate dimensions
```{r}
tiff("C:/github/WBR-microbiome/figures/CAZy_presence_dotchart_filtered.tiff", width = 8.4, height = 6.5, units = 'in', res = 150)
plot(p2) # Make plot
dev.off()
```